Synthesising Termination Proofs for Bit-Vector Programs

### 2014-07-06 papers/termination/paper.tex
nudge

Somehow expected, most of the techniques are specialised in the synthesis of
linear ranking functions for linear programs over integers (or rationals)
\cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/LeeWY12,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,
DBLP:conf/atva/HeizmannHLP13,DBLP:conf/cav/BradleyMS05,DBLP:conf/tacas/CookSZ13,DBLP:conf/cav/KroeningSTW10}. 
Among them, Lee et al.  make use of transition predicate abstraction,
algorithmic learning, and decision procedures
%to compute transition invariants as proofs for the termination of linear programs 
\cite{DBLP:conf/cav/LeeWY12},
Leike and Heizmann propose linear ranking templates 
%in order to synthesise
%termination arguments for linear loop programs 
\cite{DBLP:conf/tacas/LeikeH14}, whereas
Bradley et al. compute lexicographic linear ranking functions supported by inductive linear invariants 
%for loops with linear guards and transitions: 
\cite{DBLP:conf/cav/BradleyMS05}. 
%Learning: \cite{DBLP:conf/cav/HeizmannHP14}

While the synthesis of termination arguments for linear programs over
integers is indeed well-covered in the literature, there is very limited
work for programs over machine integers.  Cook et al.  present a method
based on a reduction to Presburger arithmetic, and a template-matching
approach for predefined classes of ranking functions based on reduction to
SAT- and QBF-solving~\cite{DBLP:conf/tacas/CookKRW10}.  Similarly, the only
work we are aware of that can compute non-linear ranking functions for
imperative loops with polynomial guards and polynomial assignments
is~\cite{DBLP:conf/vmcai/BradleyMS05}.  However, this work extends only to
polynomials.

@@

As mentioned in Section~\ref{sec:intro}, approaches based on Ramsey's
theorem compute a set of local termination conditions that decrease as
execution follows through the loop and require expensive reachability
analyses~\cite{DBLP:conf/lpe/CodishG03,DBLP:conf/lics/PodelskiR04,DBLP:conf/pldi/CookPR06}. 
In an attempt to reduce the complexity of checking the validity of the
termination argument, Cook et al present an iterative termination proving
procedure that searches for lexicographic termination
arguments~\cite{DBLP:conf/tacas/CookSZ13}, whereas Kroening et al. 
strengthen the termination argument such that it becomes a transitive
relation~\cite{DBLP:conf/cav/KroeningSTW10}.

Proving program termination implies the simultaneous search for a
termination argument and a supporting invariant.  
%For this purpose, several tools alternate between calls to a safety prover and calls to a ranking function synthesis tool \cite{}.  
%In order to share more information about
%the state of the termination proofs between the two tools, 
Brockschmidt et al. 
share the same representation of the state of the termination proof
between the safety prover and the ranking function synthesis tool \cite{DBLP:conf/cav/BrockschmidtCF13}.  
In \cite{DBLP:conf/cav/BradleyMS05},
Bradley et al.  combine the generation of ranking functions with the
generation of invariants to form one constraint solving problem such that
the necessary supporting invariants for the ranking function are discovered
on demand.  In our setting, both the ranking function and the supporting
invariant are iteratively constructed through the same refinement loop.

While program termination has been extensively studied, much less research has been conducted 
in the area of proving non-termination. Gupta et al. dynamically enumerate lasso-shaped candidate
paths for counterexamples, and then statically prove their feasibility \cite{DBLP:conf/popl/GuptaHMRX08}. 
In \cite{DBLP:conf/tacas/ChenCFNO14}, Chen et al. prove nontermination via
reduction to safety proving. Their iterative algorithm uses counterexamples to a fixed safety property
to refine an underapproximation of a program.

>>>>

Somehow expected, most of the techniques are specialised in the synthesis of
linear ranking functions for linear programs over integers (or rationals)
\cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/LeeWY12,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,
DBLP:conf/atva/HeizmannHLP13,DBLP:conf/cav/BradleyMS05,DBLP:conf/tacas/CookSZ13,DBLP:conf/cav/KroeningSTW10}. 
Among them, Lee et al.~make use of transition predicate abstraction,
algorithmic learning, and decision procedures
%to compute transition invariants as proofs for the termination of linear programs 
\cite{DBLP:conf/cav/LeeWY12},
Leike and Heizmann propose linear ranking templates 
%in order to synthesise
%termination arguments for linear loop programs 
\cite{DBLP:conf/tacas/LeikeH14}, whereas
Bradley et al.~compute lexicographic linear ranking functions supported by inductive linear invariants 
%for loops with linear guards and transitions: 
\cite{DBLP:conf/cav/BradleyMS05}. 
%Learning: \cite{DBLP:conf/cav/HeizmannHP14}

While the synthesis of termination arguments for linear programs over
integers is indeed well-covered in the literature, there is very limited
work for programs over machine integers.  Cook et al.~present a method
based on a reduction to Presburger arithmetic, and a template-matching
approach for predefined classes of ranking functions based on reduction to
SAT- and QBF-solving~\cite{DBLP:conf/tacas/CookKRW10}.  Similarly, the only
work we are aware of that can compute non-linear ranking functions for
imperative loops with polynomial guards and polynomial assignments
is~\cite{DBLP:conf/vmcai/BradleyMS05}.  However, this work extends only to
polynomials.

@@

As mentioned in Section~\ref{sec:intro}, approaches based on Ramsey's
theorem compute a set of local termination conditions that decrease as
execution follows through the loop and require expensive reachability
analyses~\cite{DBLP:conf/lpe/CodishG03,DBLP:conf/lics/PodelskiR04,DBLP:conf/pldi/CookPR06}. 
In an attempt to reduce the complexity of checking the validity of the
termination argument, Cook et al.~present an iterative termination proving
procedure that searches for lexicographic termination
arguments~\cite{DBLP:conf/tacas/CookSZ13}, whereas Kroening et
al.~strengthen the termination argument such that it becomes a transitive
relation~\cite{DBLP:conf/cav/KroeningSTW10}.

Proving program termination implies the simultaneous search for a
termination argument and a supporting invariant.  
%For this purpose, several tools alternate between calls to a safety prover and calls to a ranking function synthesis tool \cite{}.  
%In order to share more information about
%the state of the termination proofs between the two tools, 
Brock\-schmidt et al.~share the same representation of the state of the termination proof
between the safety prover and the ranking function synthesis tool \cite{DBLP:conf/cav/BrockschmidtCF13}.  
In \cite{DBLP:conf/cav/BradleyMS05},
Bradley et al.~combine the generation of ranking functions with the
generation of invariants to form one constraint solving problem such that
the necessary supporting invariants for the ranking function are discovered
on demand.  In our setting, both the ranking function and the supporting
invariant are iteratively constructed through the same refinement loop.

While program termination has been extensively studied, much less research
has been conducted in the area of proving non-termination.  Gupta et
al.~dynamically enumerate lasso-shaped candidate paths for counterexamples,
and then statically prove their
feasibility~\cite{DBLP:conf/popl/GuptaHMRX08}. 
In~\cite{DBLP:conf/tacas/ChenCFNO14}, Chen et al.~prove nontermination via
reduction to safety proving.  Their iterative algorithm uses counterexamples
to a fixed safety property to refine an underapproximation of a program.
